---
title: "Assignment Week2"
author: "Ankit Rautela"
date: "August 6, 2016"
output: md_document
---

## 1. Loading the data


```{r dataload, echo=TRUE}

activity<-read.csv("activity.csv")
head(activity)
```

## 2. Mean total number of steps taken per day

### Total Steps:

```{r mean, echo=TRUE}
totalsteps<-aggregate(activity$steps,list(activity$date),sum)
names(totalsteps)<-c("Date", "Total Steps")
head(totalsteps)
```
### Histogram for total steps per day:

```{r histogram, echo=TRUE}
hist(totalsteps$`Total Steps`,main="Total steps per day", xlab="Steps")
```

### mean and median of the total number of steps taken per day
 
 
```{r summary, echo=TRUE}
summary(totalsteps)
```

## 3. The average daily activity pattern


```{r meaninterval, echo=TRUE}

meaninterval<-aggregate(activity$steps,list(activity$interval),mean,na.rm=TRUE)
names(meaninterval)<-c("Interval","Mean")
plot(meaninterval$Interval,meaninterval$Mean,type = "l",main="Average daily activity pattern", xlab="Interval",ylab="Mean Steps")
```
    
    
### Maximum number of steps in an interval
     

```{r maxsteps }
subset(meaninterval$Interval,meaninterval$Mean==max(meaninterval$Mean))
```





## 3. Imputing missing values

### Calculate the number of missing values
```{r Na, echo=TRUE}

sum(is.na(activity))
```

### Filling the missing values

```{r filling, echo=TRUE}
activityna<-activity
i<-nrow(activity)

for (j in 1:i)
{
  if(is.na(activityna$steps[j]))
  {
    k<-activityna$interval[j]
    
    activityna$steps[j]<-subset(meaninterval$Mean,meaninterval$Interval==k)
  }
  
  
}
```

### histogram of the total number of steps taken each day
```{r histogramnew, echo=TRUE}
totalstepsnew<-aggregate(activityna$steps, list(activityna$date),sum)
names(totalstepsnew)<-c("Date","Total Steps")
hist(totalstepsnew$`Total Steps`,main="Total steps per day", xlab="Steps")
```


### New Mean and Median

```{r meannew, echo=TRUE}
mean(totalstepsnew$`Total Steps`)
median(totalstepsnew$`Total Steps`)
```

### Old Mean and Median

```{r meanold, echo=TRUE}
mean(totalsteps$`Total Steps`,na.rm=TRUE)
median(totalsteps$`Total Steps`,na.rm=TRUE)
```

Median values are slightly different and mean values are exactly the same

## 4.Calculate Weekdays and Weekends

```{r weeks, echo=TRUE}
week<-0
for (i in 1:nrow(activityna))
{
  week[i]<-weekdays(as.Date(activityna$date[i]))
  if(week[i] %in% c("Saturday","Sunday") )
  {
    week[i]<-"weekend"
  }
  else {
    week[i]<- "weekday"
  }
    
  
}
activitynew<-cbind(activityna,week)

```
### Calculate the aggregation

```{r agg, echo=TRUE}
meanstepsweek<-aggregate(activitynew$steps, list(activitynew$week,activitynew$interval),mean)
names(meanstepsweek)<-c("Week","Interval","Steps")

```




### Plot data on weekdays vs weekends

```{r plot, echo=TRUE}
library(ggplot2)
g<-ggplot(meanstepsweek,aes(Interval,Steps))
g+geom_line()+facet_wrap(~Week,ncol=1)
```